---
title: "Royal Society Corpus: How Metadata Enables Different Research Questions"
author: "PhD Workshop Example"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      cache = TRUE, fig.width = 10, fig.height = 6)
```

# Introduction: Metadata as Research Infrastructure

This analysis demonstrates how different types of metadata enable fundamentally different research questions in text analysis. We'll use the Royal Society Corpus to show how temporal, network, topical, and authorship metadata each unlock new analytical possibilities.

**Key Learning Goal**: Understanding that your research questions are fundamentally shaped and limited by the metadata available in your corpus.

## Load libraries and data

The data below was constructed using load_and_prepare_data.Rmd

It includes:
- rsc_corpus: full corpus with metadata
- network_data: separate network/community dataset
- subset_corpus(): function for flexible subsetting

```{r libraries}
library(quanteda)
library(quanteda.textplots)
library(quanteda.textstats)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(purrr)
library(DT)
library(plotly)

custom_stopwords <- c("fig", "page", "pp", "vol", "etc", "viz", "mr", "dr", "prof", "#151",
                  "#183", "also", "upon", "one", "two", "three", "four", "five", "six", "seven",
                  "eight", "nine", "ten", "may", "can", "+", "-", "~", "=", "^",
                  "per" ,"cent", "royal", "society", "phil", "trans", "8vo", "|", "results",
                  "another", "professor",
                  letters)
```


```{r load-dependencies}
# Load RSC corpus
rsc_corpus <- readRDS("../data/rsc_corpus.RDS")

# Load network/community data
network_data <- read.csv("../data/rsc/multi_temporal_communities.csv")

cat("Network data dimensions:", nrow(network_data), "rows x", ncol(network_data), "columns\n")
cat("Authors in network data:", sum(network_data$node_type == "author", na.rm = TRUE), "\n")

# Load helper functions for subsetting/analysing data
source("workshop_functions.R")
```

# Data overviews

 "slice_1650_1699"        
 [5] "slice_1675_1724"         "slice_1700_1749"         "slice_1725_1774"         "slice_1750_1799"        
 [9] "slice_1775_1824"         "slice_1800_1849"         "slice_1825_1874"         "slice_1850_1899"        
[13] "slice_1875_1924"      

```{r network_overview}
# Netowrk/community data
cat("Network data dimensions:", nrow(network_data), "rows x", ncol(network_data), "columns\n")
cat("Authors in network data:", sum(network_data$node_type == "author", na.rm = TRUE), "\n")

# Show available communities
slice_cols <- grep("slice_1", names(network_data))
for(slice_col in slice_cols) {
  cat("Time slice for:", names(network_data[slice_col]), "\n")
  slice_communities <- unique(network_data[,slice_col])
  slice_communities <- slice_communities[-c(which(is.na(slice_communities)))] # Remove NAs
  slice_communities <- slice_communities[-c(which(slice_communities == 0))] # Remove 0 which represents isolates
  cat("Total communities for timeslice:", length(slice_communities), "\n")
}

# Communities by entire dataset
per_time_slicecommunities <- unique(network_data[!is.na(network_data)])
cat("Available communities (for the entire time period):", length(all_time_communities), "\n")

# Example: Get authors from community 1
comm1_authors <- get_authors_from_communities(1, "all_time_communities")
cat("Authors in community 1:", length(comm1_authors), "\n")
if (length(comm1_authors) > 0) {
  cat("First few authors:", paste(head(comm1_authors, 3), collapse = ", "), "\n")
}

# Example: Get community info for specific authors
example_authors <- c("Sir William Ramsay, K. C. B., F. R. S.", "Professor C. Niven, F. R. S.")
author_comms <- get_author_communities(example_authors, "all_time_communities")
if (nrow(author_comms) > 0) {
  cat("Community memberships:\n")
  print(author_comms)
}

p6 <- network_data %>%
  filter(node_type == "author", !is.na(all_time_communities), all_time_communities != 0) %>%
  count(all_time_communities) %>%
  ggplot(aes(x = reorder(factor(all_time_communities), -n), y = n)) +
  geom_col(fill = "red") +
  labs(title = "Authors by Community (All-Time)", 
       x = "Community", y = "Number of Authors") +
  theme_minimal()

print(p6)
```

```{r corpus_overview}
cat("Total documents in corpus:", ndoc(rsc_corpus), "\n")
cat("Corpus docvar names:", paste(names(docvars(rsc_corpus)), collapse = ", "), "\n\n")


```

## Authors

```{r author_stats}

# Display author summary stats

cat("Documents with authors:", sum(docvars(rsc_corpus)$author_count > 0), "\n")
cat("Single-author documents:", sum(docvars(rsc_corpus)$author_count == 1), "\n")
cat("Multi-author documents:", sum(docvars(rsc_corpus)$author_count > 1), "\n")
cat("Max authors per document:", max(docvars(rsc_corpus)$author_count), "\n")
cat("Min authors per document:", min(docvars(rsc_corpus)$author_count), "\n") # Some documents have no author infomration

# Documents by authorship
p4 <- docvars(rsc_corpus)$author_count %>%
  data.frame(author_count = .) %>%
  count(author_count) %>%
  ggplot(aes(x = author_count, y = n)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(title = "Distribution of Author Count per Document",
       x = "Number of Authors", y = "Number of Documents") +
  theme_minimal()

p5 <- data.frame(
  decade = docvars(rsc_corpus)$decade,
  is_multi_author = docvars(rsc_corpus)$is_multi_author
) %>%
  group_by(decade, is_multi_author) %>%
  summarise(count = n(), .groups = "drop") %>%
  ggplot(aes(x = decade, y = count, fill = is_multi_author)) +
  geom_col(position = "dodge", alpha = 0.7) +
  labs(
    title = "Single vs Multi-Author Documents by Decade", 
    x = "Decade", y = "Number of Documents",
    fill = "Multi-Author"
  ) +
  theme_minimal()

print(p4)
print(p5)
```

## Topics


```{r topic_overview}
# Check topics
topic_values <- docvars(rsc_corpus, "primaryTopic")
cat("Unique primary topics (first 10):", paste(sort(unique(topic_values))[1:10], collapse = ", "), "\n")
# physics_topics <- c("Atomic Physics", "Thermodynamics", "Fluid Dynamics")
# cat("Physics topics found:", paste(physics_topics[physics_topics %in% unique(topic_values)], collapse = ", "), "\n")
# cat("Number of physics docs:", sum(topic_values %in% physics_topics, na.rm = TRUE), "\n\n")

secondary_topic_values <- docvars(rsc_corpus, "secondaryTopic")
cat("Unique secondary topics (first 10):", paste(sort(unique(secondart_topic_values))[1:10], collapse = ", "), "\n")

# Combine counts for both topic types
topic_counts <- df_complete %>%
  data.frame(primaryTopic = topic_values,
             secondaryTopic = secondary_topic_values) %>%
  tidyr::pivot_longer(cols = c(primaryTopic, secondaryTopic),
                      names_to = "TopicType", values_to = "Topic") %>%
  count(TopicType, Topic, sort = TRUE)


p3 <- topic_counts %>%
  ggplot(aes(x = reorder(Topic, n), y = n, fill = TopicType)) +
  geom_col(position = position_dodge()) +
  coord_flip() +
  labs(title = "Top Primary and Secondary Topics", 
       x = "Topic", y = "Number of Documents", fill = "Topic Type") +
  theme_minimal()

print(p3)
```

## Years

```{r time_overview}
# Check decade values
decade_values <- docvars(rsc_corpus, "decade")
cat("Unique decade values:", paste(sort(unique(decade_values)), collapse = ", "), "\n")

# Documents by decade
p1 <- decade_values %>%
  data.frame(decade = decade_values) %>%
  ggplot(aes(x = decade)) +
  geom_bar(fill = "steelblue", alpha = 0.7) +
  labs(title = "Documents by Decade", 
       x = "Decade", y = "Number of Documents") +
  theme_minimal()

print(p1)
```

## Publication types

```{r Publication_type_overviews}
# Check document types
type_values <- docvars(rsc_corpus, "type")
cat("Unique document types:", paste(unique(type_values), collapse = ", "), "\n")
cat("Number of article docs:", sum(type_values == "article", na.rm = TRUE), "\n\n")

# Documents by type
p2 <- type_values %>%
  data.frame(type = type_values) %>%
  count(type) %>%
  ggplot(aes(x = reorder(type, n), y = n)) +
  geom_col(fill = "forestgreen", alpha = 0.7) +
  coord_flip() +
  labs(title = "Documents by Type", 
       x = "Document Type", y = "Number of Documents") +
  theme_minimal()

print(p2)
```

# Test Individual Subsetting Functions

EXPLAIN HOW FUNCTION WORKS

```{r test-individual-functions}
# Test date subsetting directly
test_1900s <- subset_by_date(rsc_corpus, decade = 1900)
cat("Direct decade subsetting (1900s):", ndoc(test_1900s), "documents\n")

# Test type subsetting directly
test_articles <- subset_by_type(rsc_corpus, "article")
cat("Direct type subsetting (articles):", ndoc(test_articles), "documents\n")

# Test topic subsetting directly
test_physics <- subset_by_topic(rsc_corpus, physics_topics)
cat("Direct topic subsetting (physics):", ndoc(test_physics), "documents\n")

# Test author count subsetting directly
test_multi_author <- subset_by_author_count(rsc_corpus, multi_author = TRUE)
cat("Direct multi-author subsetting:", ndoc(test_multi_author), "documents\n")
```
### Example Subsetting Operations

```{r subsetting-examples}
# Example 1: Subset by decade (1900s) with debugging to show how filtering is working

corpus_1900s <- subset_corpus(rsc_corpus, decade = 1900, debug = TRUE)
cat("1900s corpus:", ndoc(corpus_1900s), "documents\n\n")

# Example 2: Subset by document type (articles only) with debugging
corpus_articles <- subset_corpus(rsc_corpus, doc_types = "article", debug = TRUE)
cat("Articles only:", ndoc(corpus_articles), "documents\n\n")

# Example 3: Multi-author documents with debugging
corpus_multi_author <- subset_corpus(rsc_corpus, multi_author = TRUE, debug = TRUE)
cat("Multi-author documents:", ndoc(corpus_multi_author), "documents\n\n")

# Let's also test the other examples without debug for cleaner output
# Example 4: Subset by topic (Physics-related)
physics_topics <- c("Atomic Physics", "Thermodynamics", "Fluid Dynamics")
corpus_physics <- subset_corpus(rsc_corpus, topics = physics_topics)
cat("Physics topics:", ndoc(corpus_physics), "documents\n")

# Example 5: Combined subsetting (1900s physics articles)
corpus_1900s_physics <- subset_corpus(rsc_corpus, 
                                     decade = 1900, 
                                     doc_types = "article",
                                     topics = physics_topics)
cat("1900s physics articles:", ndoc(corpus_1900s_physics), "documents\n")

# Example 6: Documents by specific authors
example_authors <- c("Sir William Ramsay, K. C. B., F. R. S.", "Professor C. Niven, F. R. S.")
corpus_by_authors <- subset_corpus(rsc_corpus, authors = example_authors)
cat("Documents by example authors:", ndoc(corpus_by_authors), "documents\n")

# Example 7: Documents from specific communities (using separate network data)
example_communities <- c(1, 2)
corpus_communities_alltime <- subset_corpus(rsc_corpus, communities = example_communities)
cat("Documents from communities 1 & 2 (all-time):", ndoc(corpus_communities_alltime), "documents\n")

# Example 8: Community analysis for specific time period
corpus_communities_1900s <- subset_corpus(rsc_corpus, communities = c(1, 2), 
                                         time_slice = "slice_1875_1924")
cat("Documents from communities 1 & 2 (1875-1924 slice):", ndoc(corpus_communities_1900s), "documents\n")

# Example 9: Get authors from specific communities first, then analyze
authors_comm1 <- get_authors_from_communities(communities = 1, time_slice = "all_time_communities")
cat("Authors in community 1:", length(authors_comm1), "authors\n")
if (length(authors_comm1) > 0) {
  corpus_comm1_authors <- subset_corpus(rsc_corpus, authors = authors_comm1[1:min(3, length(authors_comm1))])
  cat("Documents by first 3 authors from community 1:", ndoc(corpus_comm1_authors), "documents\n")
}
```

# 1. Baseline: Text Without Metadata Context

**Research Questions Possible**: What are the most common concepts discussed by the Royal Society overall?



```{r baseline-analysis}

# Create basic corpus for analysis (full dataset)
full_corpus <- rsc_corpus

# Create tokens and remove stopwords
full_tokens <- tokens(full_corpus, remove_punct = TRUE, remove_numbers = TRUE, 
                      remove_symbols = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

# Document-feature matrix
full_dfm <- dfm(full_tokens)

# Most frequent terms
cat("Top 20 Most Frequent Terms (All Documents):\n")
topfeatures(full_dfm, 20) %>% print()

# Most frequent bigrams
full_bigrams <- tokens_ngrams(full_tokens, n = 2) %>% dfm()
cat("\nTop 10 Most Frequent Bigrams:\n")
topfeatures(full_bigrams, 10) %>% print()

cat("\n" , rep("=", 60), "\n")
```

# 2. Temporal Metadata: Historical Evolution

**Research Questions Enabled**: How did scientific concepts evolve over time? When did new terminology emerge?

```{r temporal-analysis}
cat("=== ANALYSIS 2: TEMPORAL COMPARISON (1700s vs 1800s) ===\n")
cat("Question: How did scientific vocabulary change between the 19th and 20th centuries?\n\n")

# Create temporal subsets
corpus_1700s <- subset_corpus(rsc_corpus, start_year = 1700, end_year = 1799)
corpus_1800s <- subset_corpus(rsc_corpus, start_year = 1800, end_year = 1899)

# Overview of temporal subsets
cat("\n1700s Corpus:\n") 
cat("Documents:", ndoc(corpus_1700s), "\n")
cat("Tokens:", sum(ntoken(corpus_1700s)), "\n\n")

cat("\n1800s Corpus:\n")
cat("Documents:", ndoc(corpus_1800s), "\n")
cat("Tokens:", sum(ntoken(corpus_1800s)), "\n")

# Process tokens for both periods
tokens_1800s <- tokens(corpus_1800s, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

tokens_1700s <- tokens(corpus_1700s, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

# Create DFMs
dfm_1800s <- dfm(tokens_1800s)
dfm_1700s <- dfm(tokens_1700s)

# Compare top terms
cat("\nTop 15 Terms in 1700s:\n")
topfeatures(dfm_1700s, 15) %>% print()

cat("Top 15 Terms in 1800s:\n")
topfeatures(dfm_1800s, 15) %>% print()

# Look for terms that became more prominent in 1800s
# Simple frequency comparison (normalized by corpus size)
freq_1700s <- textstat_frequency(dfm_1700s) %>%
  mutate(rel_freq_1700s = frequency / sum(ntoken(corpus_1700s))) %>%
  select(feature, rel_freq_1700s)

freq_1800s <- textstat_frequency(dfm_1800s) %>% 
  mutate(rel_freq_1800s = frequency / sum(ntoken(corpus_1800s))) %>%
  select(feature, rel_freq_1800s)

# Find terms that increased dramatically
freq_comparison <- merge(freq_1700s, freq_1800s, by = "feature", all = TRUE)
freq_comparison[is.na(freq_comparison)] <- 0
freq_comparison$ratio <- freq_comparison$rel_freq_1800s / (freq_comparison$rel_freq_1700s + 0.0001)

cat("\nTerms that became much more frequent in 1800s (ratio > 3):\n")
freq_comparison %>%
  filter(rel_freq_1700s > 0.0001, ratio > 3) %>%
  arrange(desc(ratio)) %>%
  head(10)

top_emerging_word <- freq_comparison %>%
    filter(rel_freq_1700s > 0.0001, ratio > 3) %>%
    arrange(desc(ratio)) %>%
    head(1)

top_emerging_word <- top_emerging_word$feature

# KWIC analysis for an emerging term
if(sum(stringr::str_detect(featnames(dfm_1800s), "electron")) > 0) {
  cat("\nKeyword in Context:", top_emerging_word, "in 1800s texts:\n")
  kwic(tokens_1800s, top_emerging_word, window = 5) %>% head(5) 
}

cat("\n" , rep("=", 60), "\n")
```

# 3. Network/Community Metadata: Intellectual Communities

**Research Questions Enabled**: How do different scholarly communities use language differently? What concepts are community-specific vs. shared?

```{r network-analysis}
cat("=== ANALYSIS 3: COMMUNITY COMPARISON ===\n")
cat("Question: How do different intellectual communities discuss scientific concepts?\n\n")

# NB: Community zero represents isolates - needs to be dropped!
# Get the largest communities for meaningful comparison
large_communities <- network_data %>%
  filter(all_time_communities != 0) %>%
  filter(node_type == "author", !is.na(all_time_communities)) %>%
  count(all_time_communities, sort = TRUE) %>%
  head(3) %>%
  pull(all_time_communities)

cat("Analyzing communities:", paste(large_communities, collapse = ", "), "\n\n")

# Create community-based subsets
corpus_comm1 <- subset_corpus(rsc_corpus, communities = large_communities[1])
corpus_comm2 <- subset_corpus(rsc_corpus, communities = large_communities[2])

# Overview of community subsets
cat("Community", large_communities[1], "Corpus:\n")
cat("Documents:", ndoc(corpus_comm1), "\n")
cat("Tokens:", sum(ntoken(corpus_comm1)), "\n")

cat("\nCommunity", large_communities[2], "Corpus:\n")
cat("Documents:", ndoc(corpus_comm2), "\n") 
cat("Tokens:", sum(ntoken(corpus_comm2)), "\n\n")

if(ndoc(corpus_comm1) > 0 && ndoc(corpus_comm2) > 0) {
  # Process tokens for both communities
  tokens_comm1 <- tokens(corpus_comm1, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_tolower() %>%
    tokens_remove(stopwords("english")) %>%
    tokens_remove(custom_stopwords)
  
  tokens_comm2 <- tokens(corpus_comm2, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_tolower() %>%
    tokens_remove(stopwords("english")) %>%
    tokens_remove(custom_stopwords)
  
  # Create DFMs
  dfm_comm1 <- dfm(tokens_comm1)
  dfm_comm2 <- dfm(tokens_comm2)
  
  # Compare vocabularies
  cat("Top 15 Terms in Community", large_communities[1], ":\n")
  topfeatures(dfm_comm1, 15) %>% print()
  
  cat("\nTop 15 Terms in Community", large_communities[2], ":\n")
  topfeatures(dfm_comm2, 15) %>% print()
  
  # Find community-distinctive terms using keyness
  combined_dfm <- rbind(dfm_comm1, dfm_comm2)
  docvars(combined_dfm, "community") <- c(rep(paste("Community", large_communities[1]), ndoc(dfm_comm1)),
                                          rep(paste("Community", large_communities[2]), ndoc(dfm_comm2)))
  
  # Calculate keyness
  keyness_results <- textstat_keyness(combined_dfm, 
                                      target = docvars(combined_dfm, "community") == paste("Community", large_communities[1]))
  
  cat("\nTerms most distinctive of Community", large_communities[1], ":\n")
  keyness_results %>% 
    filter(chi2 > 0) %>% 
    arrange(desc(chi2)) %>% 
    head(10) %>% 
    print()
  
  cat("\nTerms most distinctive of Community", large_communities[2], ":\n")
  keyness_results %>% 
    filter(chi2 < 0) %>% 
    arrange(chi2) %>% 
    head(10) %>% 
    print()
} else {
  cat("Insufficient documents in one or both communities for comparison.\n")
}

cat("\n" , rep("=", 60), "\n")
```

# 4. Topical Metadata: Disciplinary Language

**Research Questions Enabled**: How do different fields use scientific language? What terminology is field-specific?

```{r topic-analysis}
cat("=== ANALYSIS 4: TOPICAL COMPARISON ===\n")
cat("Question: How does scientific language differ between research areas?\n\n")

# Find the most common topics for comparison
common_topics <- df_complete %>%
  count(primaryTopic, sort = TRUE) %>%
  filter(!is.na(primaryTopic)) %>%
  head(4) %>%
  pull(primaryTopic)

cat("Analyzing topics:", paste(common_topics[1:2], collapse = " vs "), "\n\n")

# Create topic-based subsets (using first two most common topics)
corpus_topic1 <- subset_corpus(rsc_corpus, topics = common_topics[1])
corpus_topic2 <- subset_corpus(rsc_corpus, topics = common_topics[2])

# Overview of topic subsets
cat(common_topics[1], "Corpus:\n")
cat("Documents:", ndoc(corpus_topic1), "\n")
cat("Tokens:", sum(ntoken(corpus_topic1)), "\n")

cat("\n", common_topics[2], "Corpus:\n")
cat("Documents:", ndoc(corpus_topic2), "\n")
cat("Tokens:", sum(ntoken(corpus_topic2)), "\n\n")

# Process tokens for both topics
tokens_topic1 <- tokens(corpus_topic1, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

tokens_topic2 <- tokens(corpus_topic2, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

# Create DFMs
dfm_topic1 <- dfm(tokens_topic1)
dfm_topic2 <- dfm(tokens_topic2)

# Compare vocabularies
cat("Top 15 Terms in", common_topics[1], ":\n")
topfeatures(dfm_topic1, 15) %>% print()

cat("\nTop 15 Terms in", common_topics[2], ":\n")
topfeatures(dfm_topic2, 15) %>% print()

# Calculate distinctive terms
combined_topic_dfm <- rbind(dfm_topic1, dfm_topic2)
docvars(combined_topic_dfm, "topic") <- c(rep(common_topics[1], ndoc(dfm_topic1)),
                                          rep(common_topics[2], ndoc(dfm_topic2)))

topic_keyness <- textstat_keyness(combined_topic_dfm, 
                                  target = docvars(combined_topic_dfm, "topic") == common_topics[1])

cat("\nTerms most distinctive of", common_topics[1], ":\n")
topic_keyness %>% 
  filter(chi2 > 0) %>% 
  arrange(desc(chi2)) %>% 
  head(10) %>% 
  print()

cat("\nTerms most distinctive of", common_topics[2], ":\n")
topic_keyness %>% 
  filter(chi2 < 0) %>% 
  arrange(chi2) %>% 
  head(10) %>% 
  print()

# Bigram analysis for topics
topic1_bigrams <- tokens_ngrams(tokens_topic1, n = 2) %>% dfm()
topic2_bigrams <- tokens_ngrams(tokens_topic2, n = 2) %>% dfm()

cat("\nTop Bigrams in", common_topics[1], ":\n")
topfeatures(topic1_bigrams, 8) %>% print()

cat("\nTop Bigrams in", common_topics[2], ":\n")
topfeatures(topic2_bigrams, 8) %>% print()

cat("\n" , rep("=", 60), "\n")
```

# 5. Authorship Metadata: Individual vs Collaborative Work

**Research Questions Enabled**: How does collaboration affect scientific language? Do single authors write differently than teams?

```{r authorship-analysis}
cat("=== ANALYSIS 5: AUTHORSHIP COMPARISON (Single vs Multi-Author) ===\n")
cat("Question: How does scientific language differ between individual and collaborative work?\n\n")

# Create authorship-based subsets
corpus_single <- subset_corpus(rsc_corpus, single_author = TRUE)
corpus_multi <- subset_corpus(rsc_corpus, multi_author = TRUE)

# Overview of authorship subsets
cat("Single-Author Corpus:\n")
cat("Documents:", ndoc(corpus_single), "\n")
cat("Tokens:", sum(ntoken(corpus_single)), "\n")

cat("\nMulti-Author Corpus:\n")
cat("Documents:", ndoc(corpus_multi), "\n")
cat("Tokens:", sum(ntoken(corpus_multi)), "\n\n")

# Process tokens
tokens_single <- tokens(corpus_single, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

tokens_multi <- tokens(corpus_multi, remove_punct = TRUE, remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(custom_stopwords)

# Create DFMs
dfm_single <- dfm(tokens_single)
dfm_multi <- dfm(tokens_multi)

# Compare vocabularies
cat("Top 15 Terms in Single-Author Papers:\n")
topfeatures(dfm_single, 15) %>% print()

cat("\nTop 15 Terms in Multi-Author Papers:\n")
topfeatures(dfm_multi, 15) %>% print()

# Distinctive language
combined_auth_dfm <- rbind(dfm_single, dfm_multi)
docvars(combined_auth_dfm, "authorship") <- c(rep("Single", ndoc(dfm_single)),
                                              rep("Multi", ndoc(dfm_multi)))

auth_keyness <- textstat_keyness(combined_auth_dfm, 
                                 target = docvars(combined_auth_dfm, "authorship") == "Single")

cat("\nTerms most distinctive of Single-Author papers:\n")
auth_keyness %>% 
  filter(chi2 > 0) %>% 
  arrange(desc(chi2)) %>% 
  head(10) %>% 
  print()

cat("\nTerms most distinctive of Multi-Author papers:\n")
auth_keyness %>% 
  filter(chi2 < 0) %>% 
  arrange(chi2) %>% 
  head(10) %>% 
  print()

cat("\n" , rep("=", 60), "\n")
```

# 6. Combined Metadata: Complex Questions

**Research Questions Enabled**: How do multiple metadata dimensions interact? What happens when we layer different contextual factors?

```{r combined-analysis}
cat("=== ANALYSIS 6: COMBINED METADATA (Topic + Time) ===\n")
cat("Question: How did language in a specific field evolve over time?\n\n")

# Example: How did physics language change from 1800s to 1900s?
physics_topics <- c("Atomic Physics", "Thermodynamics", "Fluid Dynamics", "Electricity")

# Find which physics topics actually exist in our data
available_physics <- physics_topics[physics_topics %in% unique(df_complete$primaryTopic)]
if(length(available_physics) > 0) {
  target_topic <- available_physics[1]
  cat("Analyzing evolution of", target_topic, "from 1800s to 1900s\n\n")
  
  # Create combined subsets
  physics_1800s <- subset_corpus(rsc_corpus, 
                                topics = target_topic,
                                start_year = 1800, end_year = 1899)
  
  physics_1900s <- subset_corpus(rsc_corpus,
                                topics = target_topic, 
                                start_year = 1900, end_year = 1920)
  
  cat(target_topic, "in 1800s:\n")
  cat("Documents:", ndoc(physics_1800s), "\n")
  
  cat("\n", target_topic, "in 1900s:\n")
  cat("Documents:", ndoc(physics_1900s), "\n\n")
  
  if(ndoc(physics_1800s) > 10 && ndoc(physics_1900s) > 10) {
    # Process and compare
    tokens_phys_1800s <- tokens(physics_1800s, remove_punct = TRUE) %>%
      tokens_tolower() %>%
      tokens_remove(stopwords("english")) %>%
      tokens_remove(custom_stopwords)
    
    tokens_phys_1900s <- tokens(physics_1900s, remove_punct = TRUE) %>%
      tokens_tolower() %>%
      tokens_remove(stopwords("english")) %>%
      tokens_remove(custom_stopwords)
    
    dfm_phys_1800s <- dfm(tokens_phys_1800s)
    dfm_phys_1900s <- dfm(tokens_phys_1900s)
    
    cat("Top terms in", target_topic, "1800s:\n")
    topfeatures(dfm_phys_1800s, 10) %>% print()
    
    cat("\nTop terms in", target_topic, "1900s:\n")
    topfeatures(dfm_phys_1900s, 10) %>% print()
    
    # Find evolving terminology
    combined_phys_dfm <- rbind(dfm_phys_1800s, dfm_phys_1900s)
    docvars(combined_phys_dfm, "period") <- c(rep("1800s", ndoc(dfm_phys_1800s)),
                                              rep("1900s", ndoc(dfm_phys_1900s)))
    
    phys_keyness <- textstat_keyness(combined_phys_dfm,
                                     target = docvars(combined_phys_dfm, "period") == "1900s")
    
    cat("\nTerms that became more prominent in", target_topic, "by 1900s:\n")
    phys_keyness %>%
      filter(chi2 > 0) %>%
      arrange(desc(chi2)) %>%
      head(8) %>%
      print()
    
  } else {
    cat("Insufficient documents for temporal comparison within", target_topic, "\n")
  }
} else {
  cat("No physics topics found in available data\n")
}

cat("\n" , rep("=", 60), "\n")
```

# Summary: What We've Learned

```{r summary}
cat("=== METHODOLOGICAL SUMMARY ===\n\n")

cat("1. BASELINE (No Metadata): Only reveals general vocabulary patterns\n")
cat("   - Limitation: No context for interpretation\n\n")

cat("2. TEMPORAL METADATA: Enables historical analysis\n") 
cat("   - New questions: When did concepts emerge? How did language evolve?\n\n")

cat("3. NETWORK METADATA: Reveals social dimensions of knowledge\n")
cat("   - New questions: How do communities differ? What language is shared vs. specialized?\n\n")

cat("4. TOPICAL METADATA: Shows disciplinary differences\n")
cat("   - New questions: How do fields use language differently? What is field-specific?\n\n")

cat("5. AUTHORSHIP METADATA: Reveals collaborative effects\n")
cat("   - New questions: How does collaboration affect language? Individual vs. team patterns?\n\n")

cat("6. COMBINED METADATA: Enables complex, nuanced questions\n")
cat("   - New questions: How do multiple factors interact? Layered contextual analysis\n\n")

cat("KEY INSIGHT: Metadata doesn't just provide context—it fundamentally determines\n")
cat("what research questions are possible to ask and answer.\n\n")

cat("For students: Always consider what metadata dimensions are available in your\n")
cat("corpus and how they could enable different analytical approaches.\n")
```

# Student Exercise Template

```{r exercise-template, eval=FALSE}
# TEMPLATE FOR STUDENT ANALYSIS
# Students can modify these parameters to explore different questions:

# Choose your comparison:
my_decade1 <- 1850  # or any decade
my_decade2 <- 1900  # or any decade  
my_topic <- "Geology"  # or any available topic
my_community <- 1  # or any available community

# Create your subsets:
my_corpus1 <- subset_corpus(rsc_corpus, decade = my_decade1, topics = my_topic)
my_corpus2 <- subset_corpus(rsc_corpus, decade = my_decade2, topics = my_topic)

# Analyze and compare:
# [Students add their own analysis code here]

# Reflection questions for students:
# 1. What research question does your comparison address?
# 2. How did the metadata enable this specific question?
# 3. What would be impossible to answer without this metadata?
# 4. What additional metadata would enable even more interesting questions?
```